\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2018-10-22 Documents/M1/Astronum_2018/sections/Acknowledge.tex
Updated Astronum_2018.

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725.  
This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.  
This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research.
Eirik Endeve and Anthony Mezzacappa acknowledge support from the National Science Foundation under group NSF-GP-... and from the UT-ORNL Joint Institute for Computational Sciences.

>>>>

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725.  
This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.  
This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research.
Eirik Endeve and Anthony Mezzacappa acknowledge support from the National Science Foundation under group \rc{NSF-GP-... (number?)} and from the UT-ORNL Joint Institute for Computational Sciences.

### 2018-10-22 Documents/M1/Astronum_2018/sections/Conclusion.tex
Buckup the latest changes on M1/Astronum_2018
Updated Astronum_2018.

Two PD-ARS schemes are proposed for the desired IMEX scheme.
The one with SSPRK2 is second-order accurate and strong-stability preserving in the streaming limit while the other with SSPRK3 is third-order accurate.
Their properties, both accuracy and convex-invariance, were demonstrated with numerical tests.

>>>>

Two PD-ARS schemes are proposed for the desired IMEX scheme.
The one with SSPRK2 is second-order accurate and strong-stability preserving in the streaming limit while the other with SSPRK3 is third-order accurate.
Their accuracy were demonstrated with some problems with known smooth solutions in streaming, absorption, ans scattering-dominated regimes.
Neutrino transport test with neutrino emission, absorption, elastic scattering, and stationary background, was designed to test the convex-invariance of our PD-ARS schemes.
The neutrino stationary state test shows that it's necessary to combine an algebraic closure based on Fermi-Dirac statistics and convex-invariant time integration for a solid CCSN simulation.

### 2018-10-22 Documents/M1/Astronum_2018/sections/Numerical.tex
Updated Astronum_2018.

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution, adopted from \cite{radice_etal_2013}, is the sine wave diffusion test, i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain with $D=\{x:x\in[-3,3]\}$ and an initial condition with $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ is used.
The analytic solution is given by $\cJ = \cJ_{0} \times e^{-\f{\pi^2 t}{27\sigma_{\Scatt}}}$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$. 
Then the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the error versus $N$.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
PD-ARS with SSPRK2 behaves as good as SSP2332 in the diffusion region but requires 1/3 less implicit solvers.
\begin{figure}[h]
  \centering
  \centerline{\includegraphics[width=1.2\textwidth]{figures/SineWaveDiffusion}}
%  \begin{tabular}{cc}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionJ}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionH}
%  \end{tabular}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles), and PD-ARS with SSPRK3 (hexagram). Black dashed lines on the left plot reference lines proportional to $N^{-2}$ (top) and $N^{-3}$ (bottom), respectively. Black dashed line on the right plot reference lines proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

\subsection{Neutrino Stationary State Test} \label{se: Neutrino Stationary State Test}
Next we consider a more ``realistic'' test: two-dimensional neutrino transport with emission, absorption, and elastic scattering through a stationary background.
This is designed to test the realizability-preserving properties of the PD-ARS schemes.
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities we used.
This test is constructed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction.
We take 10 energy groups from 0~MeV to 300~MeV and initialize the neutrino number density $\cJ$ and flux density $\bcH$ both at zero(arbitrarily small).
For the realizability-preserving test, we employ the CB closure with PD-ARS schemes, SSP2332, a scheme proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and a scheme proposed by McClarren et al \cite{mcclarren_etal_2008} are all applied.
Only PD-ARS schemes produce realizability-preserving results and evolve the model to a steady state.
The other schemes crashed immediately after several steps.
The realizability-preserving plot and the evolve process of PD-ARS schemes are given in Figure~\ref{fig:NeutrinoStationaryTestEvolve}.

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
    \caption{Realizability (left column), the number density $\cJ$ versus radius (center column), and the flux factor $|\bcH|/\cJ$ versus radius (right column) at t = 0.01~ms ,0.35~ms and 5.0~ms for the neutrino stationary state test. For the realizability plots, the light blue area highlights the realizability domain, and the black lines define its boundary. Each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot. The results of PD-ARS with SSPRK2 and PD-ARS with SSPRK3 are indistinguishable in these plots.}
    \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}

>>>>

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution, adopted from \cite{radice_etal_2013}, is the sine wave diffusion test, i.e. $\sigma_{\Ab} = 0$ and $f_0 = 0$.
A periodic domain with $D=\{x:x\in[-3,3]\}$ and an initial condition with $\cJ_{0} = 0.5+0.49\times\sin\big(\f{\pi\,x}{3}\big)$ and $\cH_{0} =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ_{0}}{x}$ is used.
The analytic solution is given by $\cJ = \cJ_{0} \times e^{-\f{\pi^2 t}{27\sigma_{\Scatt}}}$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$. 
Then the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the error versus $N$.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
PD-ARS with SSPRK2 behaves as good as SSP2332 in the diffusion region but requires 1/3 less implicit solvers.

\begin{figure}[h]
  \centering
  \centerline{\includegraphics[width=1.2\textwidth]{figures/SineWaveDiffusion}}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles), and PD-ARS with SSPRK3 (hexagram). Black dashed lines on the left plot reference lines proportional to $N^{-2}$ (top) and $N^{-3}$ (bottom), respectively. Black dashed line on the right plot reference lines proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

\subsection{Neutrino Stationary State Test} \label{se: Neutrino Stationary State Test}
Next we consider a more ``realistic'' test: two-dimensional neutrino transport with emission, absorption, and elastic scattering through a stationary background.
This is designed to test the realizability-preserving properties of the PD-ARS schemes.
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities, which are adopted from \cite{Bruenn_1985}, we used.
This test is constructed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction.
We take 10 energy groups from 0~MeV to 300~MeV and initialize the neutrino number density $\cJ = 10^{-99}$ and flux density $\bcH=0$.
For the realizability-preserving test, we employ the CB closure with PD-ARS schemes, SSP2332, a scheme proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and a scheme proposed by McClarren et al \cite{mcclarren_etal_2008} are all applied.
Only PD-ARS schemes produce realizability-preserving results and evolve the model to a steady state.
The other schemes crashed immediately after several steps.
The realizability-preserving plots are given in the first column of Figure~\ref{fig:NeutrinoStationaryTestEvolve} at $t=0.01$~ms, 0.35~ms and 5.0~ms.
With each states be marked as a red dot in $\cJ$-$\bcH$ space and the realizable domain be light blue region, they show that all the states achieved in the simulation with PD-ARS schemes are realizable.
The second and third column of Figure~\ref{fig:NeutrinoStationaryTestEvolve} plot the corresponding the neutrino number density $\cJ$ and flux factor $|\bcH|/\cJ$ in radius.
These plots show how the neutrinos generate and transport in the simulation.

@@

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
    \caption{Realizability (left column, light blue domain for $f \in [0,1]$ with black-solid-line as boundary, light red for $f\geq 0$ with thin-red-solid-line as boundary) , the number density $\cJ$ versus radius (center column), and the flux factor $|\bcH|/\cJ$ versus radius (right column) at t = 0.01~ms ,0.35~ms and 5.0~ms for the neutrino stationary state test. For the realizability plots, each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot, which all sites in the light blue domain. The results of PD-ARS with SSPRK2 and PD-ARS with SSPRK3 are indistinguishable in these plots.}
    \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}

### 2018-10-22 Documents/M1/Astronum_2018/sections/TimeIntegration.tex
Updated Astronum_2018.

\subsection{A Constraint-Preserving Implicit-Explicit Scheme}
Beyond the accuracy requirement, a convex-invariant IMEX scheme must satisfy additional constraints.
Our idea is finding the constraints on $a_{ij}$, $\tilde{a}_{ij}$, $\tilde{w}_{i}$, and $w_{i}$ that enable each $\vect{u}^{(i)}$ and $\vect{u}^{n+1}$ be convex combination of realizable moments:

@@

\rc{Diffusion accurate detail...}

\rc{Only ARS and define ARS...}

Here is list of our constraints \cite{chu_etal_2018}:
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item Second-order accuracy in the streaming limit:
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
      \label{eq:orderConditionsEx}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{i1} = 0, \quad \text{for} \quad i=1,\ldots,s, \nonumber \\
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    And $c_{ij}$ and $\hat{c}_{ij}$ be given by
    \begin{equation}
     \begin{aligned}
      c_{i0} &= 1-\sum_{j=2}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
      c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj} \\
      \tilde{c}_{i0} &= \tilde{a}_{i1}+\sum_{j=2}^{i-1}a_{ij}\tilde{b}_{j1}, \quad &
      \tilde{c}_{ij} &= \tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}.  
     \end{aligned}
     \label{eq:positivityCoefficientsARS}
    \end{equation}
    \begin{equation}
      b_{ii} = \f{1}{a_{ii}}, \quad
      b_{ij} = -\f{1}{a_{ii}}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad
      \tilde{b}_{ij} = -\f{1}{a_{ii}}\Big(\tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}\Big).  
    \end{equation}
    Note that $c_{i1}=\tilde{c}_{i1}=0$ in Eq.~\eqref{eq:positivityCoefficientsARS} so that $\sum_{j=0}^{i-1}c_{ij}=1$.
    \item Well-behaved in the diffusion limit:
    \begin{equation}
      \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=2,\ldots,s,
      \label{eq:diffusionCondition}
    \end{equation}
    where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix.
    \item Less than five stages ($s\le4$).
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. (GSA)
\end{enumerate}  
We call the IMEX scheme satisfying the above conditions {PD-ARS}. (Definition 3 in \cite{chu_etal_2018}.)
We give two optimized PD-ARS schemes here: PD-ARS with SSPRK2 and  PD-ARS with SSPRK3. 
They have second- and third-order accuracy in the streaming limit, respectively.
\subsubsection{PD-ARS with SSPRK2}
Here we give the optimized PD-ARS with SSPRK2 in the standard double Butcher tableau form: explicit tableau on the left ($\tilde{A}$), implicit tableau on the right ($A$):
\begin{equation}
  \begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2 & 1/2 \\ \hline
  	  & 0 & 1/2 & 1/2
  \end{array}
\end{equation}
For this scheme, $c_{\mbox{\tiny Sch}}= 1$ and only two implicit solver are needed for each time step.
\subsubsection{PD-ARS with SSPRK3}
Here we give the optimized PD-ARS with SSPRK3.
Its standard double Butcher tableau are (explicit tableau on the left, implicit tableau on the right)
\begin{equation}
  \begin{array}{c | c c c c}
  	    &     &     &     &  \\
  	 1  & 1   &     &     &  \\
  	1/2 & 1/4 & 1/4 &  \\
  	 1  & 1/6 & 1/6 & 2/3 &  \\ \hline
  	    & 1/6 & 1/6 & 2/3 &
  \end{array}
  \qquad
  \begin{array}{c | c c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1/2 & 0 & 1/4 & 1/4 \\ 
  	1 & 0 & 1/6 & 1/6 & 2/3\\\hline
  	  & 0 & 1/6 & 1/6 & 2/3
  \end{array}
\end{equation}
For this schemes, $c_{\mbox{\tiny Sch}}= 1$.
Three implicit solver are taken for each time step.
Since PD-ARS with SSPRK3 is as accurate as PD-ARS with SSPRK2 in collision involved region, see Section~\ref{se:NumericalTests} for the plots, it may not offer any advantage.

>>>>

\subsection{Constraint-Preserving Implicit-Explicit Schemes}
Beyond the accuracy requirement, a convex-invariant IMEX scheme must satisfy additional constraints.
Our idea is finding the constraints on $a_{ij}$, $\tilde{a}_{ij}$, $\tilde{w}_{i}$, and $w_{i}$ that enable each $\vect{u}^{(i)}$ and $\vect{u}^{n+1}$ be convex combination of realizable moments:

@@

\subsection{Diffusion Accurate, Constraint-Preserving Implicit-Explicit Schemes}
Diffusion accurate is the other property the time integration should have.
In the diffusion limit, the distribution function is nearly isotropic so that $\vect{\cK}\approx\f{1}{3}\,\cJ\,\vect{I}$ and $\vect{\cH}\approx-\f{1}{3}\,\tau\,\nabla\cJ$.
The moment behavior are governed by (e.g., \cite{jinLevermore_1996})
\begin{equation}
  \pd{\cJ}{t} + \nabla\cdot\vect{\cH} = 0
  \quad\text{and}\quad
  \vect{\cH} = - \tau\,\nabla\cdot\vect{\cK}.  
  \label{eq:diffusionLimit}
\end{equation}
In the context of IMEX schemes, the above relationships imply that the relation $A\,\vec{\vect{\cH}}=-\f{1}{3}\,\tau\,\tilde{A}\,\nabla\vec{\cJ}$ should hold. 
We derived the diffusion accurate condition in \cite{chu_etal_2018} and proved that only type~ARS (investigated by Ascher, Ruuth, and Spiteri\cite{ascher_etal_1997}) IMEX schemes can be our candidates.

For type~ARS, $a_{i1} = 0$ for $i=1,\ldots,s$.
$c_{ij}$ and $\hat{c}_{ij}$ are given by
    \begin{equation}
     \begin{aligned}
      c_{i0} &= 1-\sum_{j=2}^{i-1}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad &
      c_{ij} &= \sum_{l=j}^{i-1}a_{il}b_{lj}, \\
      \tilde{c}_{i0} &= \tilde{a}_{i1}+\sum_{j=2}^{i-1}a_{ij}\tilde{b}_{j1}, \quad &
      \tilde{c}_{ij} &= \tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj},  
     \end{aligned}
     \label{eq:positivityCoefficientsARS}
    \end{equation}
    \begin{equation}
      b_{ii} = \f{1}{a_{ii}}, \quad
      b_{ij} = -\f{1}{a_{ii}}\sum_{l=j}^{i-1}a_{il}b_{lj}, \quad
      \tilde{b}_{ij} = -\f{1}{a_{ii}}\Big(\tilde{a}_{ij}+\sum_{l=j+1}^{i-1}a_{il}\tilde{b}_{lj}\Big).  
    \end{equation}
    Note that $c_{i1}=\tilde{c}_{i1}=0$ in Eq.~\eqref{eq:positivityCoefficientsARS} so that $\sum_{j=0}^{i-1}c_{ij}=1$.

Combining the conditions we have, we aim to find type~ARS IMEX scheme with.
\begin{enumerate}
    \item Consistency of the implicit coefficients:
    \begin{equation}
      \sum_{i=1}^{s}w_{i}=1.
    \end{equation}
    \item Second-order accuracy in the streaming limit:
    \begin{equation}
      \sum_{i=1}^{s}\tilde{w}_{i}=1
      \quad\text{and}\quad
      \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2},
      \label{eq:orderConditionsEx}
    \end{equation}
    \item Convex-invariant:
    \begin{align}
      &a_{ii}>0, \quad c_{i0},\tilde{c}_{i0}\ge0, \quad \text{for} \quad i=2,\ldots,s, \nonumber \\
      &\text{and} \quad c_{ij},\tilde{c}_{ij}\ge0, \quad \text{for} \quad i=3,\ldots,s, \quad\text{and}\quad j=2,\ldots,i-1.  
    \end{align}
    with $\sum_{j=0}^{i-1}c_{ij}=1$, for $i=1,\ldots,s$, and $c_{\Sch}=\min_{\substack{i = 2,\ldots,s \\ 
                  j = 0,2,\ldots,i-1}}\,\f{1}{\hat{c}_{ij}}>0$.
    \item Well-behaved in the diffusion limit:
    \begin{equation}
      \vect{e}_{i}^{T}A^{-1}\tilde{A}\,\vect{e} = 1, \quad i=2,\ldots,s,
      \label{eq:diffusionCondition}
    \end{equation}
    where $\vect{e}_{i}$ is the $i$th column of the $s\times s$ identity matrix.
    \item Less than five stages ($s\le4$).
    \item Globally stiffly accurate: $a_{si}=w_{i}$ and $\tilde{a}_{si}=\tilde{w}_{i},\quad i=1,\ldots,s$. 
\end{enumerate}  
We call the IMEX scheme satisfying the above conditions {PD-ARS}. (Definition 3 in \cite{chu_etal_2018}.)
We give two optimized PD-ARS schemes here: PD-ARS with SSPRK2 and  PD-ARS with SSPRK3. 
They have second- and third-order accuracy in the streaming limit, respectively.
\subsubsection{PD-ARS with SSPRK2}
Here we give the optimized PD-ARS with SSPRK2 in the standard double Butcher tableau form: explicit tableau on the left ($\tilde{A}$), implicit tableau on the right ($A$):
\begin{equation}
  \begin{array}{c | c c c}
  	0 & 0   & 0 & 0 \\
  	1 & 1   & 0 & 0 \\
  	1 & 1/2 & 1/2 & 0 \\ \hline
  	  & 1/2 & 1/2 & 0
  \end{array}
  \qquad
  \begin{array}{c | c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1 & 0 & 1/2 & 1/2 \\ \hline
  	  & 0 & 1/2 & 1/2
  \end{array}
\end{equation}
For this scheme, $c_{\mbox{\tiny Sch}}= 1$ and only two implicit solver are needed for each time step.
\subsubsection{PD-ARS with SSPRK3}
Here we give the optimized PD-ARS with SSPRK3.
Its standard double Butcher tableau are (explicit tableau on the left, implicit tableau on the right)
\begin{equation}
  \begin{array}{c | c c c c}
  	    &     &     &     &  \\
  	 1  & 1   &     &     &  \\
  	1/2 & 1/4 & 1/4 &  \\
  	 1  & 1/6 & 1/6 & 2/3 &  \\ \hline
  	    & 1/6 & 1/6 & 2/3 &
  \end{array}
  \qquad
  \begin{array}{c | c c c c}
  	0 & 0 & 0            & 0            \\
  	1 & 0 & 1            & 0            \\
  	1/2 & 0 & 1/4 & 1/4 \\ 
  	1 & 0 & 1/6 & 1/6 & 2/3\\\hline
  	  & 0 & 1/6 & 1/6 & 2/3
  \end{array}
\end{equation}
For this schemes, $c_{\mbox{\tiny Sch}}= 1$.
Three implicit solver are taken for each time step.
Since PD-ARS with SSPRK3 is as accurate as PD-ARS with SSPRK2 in collision involved region, see Section~\ref{se:NumericalTests} for the plots, it may not offer any advantage.

### 2018-10-22 Documents/M1/Astronum_2018/sections/intro.tex
Updated Astronum_2018.

Core-collapse supernovae (CCSNe) are the explosions of massive stars that end their lives.
CCSNe are a dominant source of heavy elements and play an important role in many astrophysical phenomena, such as neutron star and black hole formation.  
Furthermore, these explosions occur at energies and densities relevant to address fundamental questions in nuclear, particle, and gravitational physics. 
A solid theoretical framework for the CCSN explosion mechanism may help answer important questions in fundamental physics.

One essential part of the explosion mechanism is neutrino transport.
The neutrino energy deposition drives the CCSN explosion.  
Ideally neutrino transport would be modeled by the Boltzmann transport equation, which is a integro-partial-differential equation evolving a phase-space distribution function $f$.
Simulating the neutrino transport implies finding a solution of the Boltzmann equation for a specific domain and period with acceptable accuracy.

However, solving the multi-dimensional Boltzmann transport equation is expensive.
To balance physical fidelity and computational expediency, an approximate method called the two-moment method has been adopted.
In the two-moment method, the evolved variables are the zeroth and first angular moments of the distribution function $f$: the spectral particle density $\cJ$ and flux $\bcH$, respectively.
However, the transport term in the Boltzmann transport equation is the dot product of the velocity vector and the gradient of the phase-space distribution function.
Integrating this term with unit direction vector as the weight introduces the second angular moment $\bcK$ to the moment equations.
Knowledge of $\bcK$ is needed to close the equation system.
Therefore, an algebraic closure that gives an approximate value of $\bcK$ given $\cJ$ and $\bcH$ is needed.
The two-moment method has been widely applied with different algebraic closures, such as the Minerbo\cite{minerbo_1978} closure(e.g.~{O'Connor} and {Couch}\cite{oConnorCouch_2018}, Pan and et al\cite{pan_etal_2018}, Glas et al\cite{glas_etal_2018}, and Just et al\cite{just_etal_2018} ) and the Levermore\cite{levermore_1984} closure(e.g.~Vartanyan et al\cite{vartanyan_etal_2018}, Cabezon et al\cite{cabezon_etal_2018}, and Kuroda et al\cite{kuroda_etal_2016}).

Applying the two-moment method alone does not guarantee an affordable solution.
To be precise, neutrinos interact with the background, and the characteristic time scale of these interactions, which can be $\sim\mathcal{O}(10^{-13})$~second, is short compared to the duration of the CCSN explosion, which $\sim\mathcal{O}(1)$~second.  
This makes the time step needed by an explicit time integrator prohibitively small. 
On the other hand, solving the moment equations fully implicitly requires inverting globally many band-structured matrices whose sizes depend on the spatial discretization.
Such a global inversion is both expensive and unfriendly to parallelization.
To circumvent these challenges, implicit-explicit (IMEX) methods are taken into consideration.
By treating the transport terms in the two-moment equations explicitly and the collision terms implicitly, IMEX methods are subject only to a time step governed by the explicit transport terms, and the matrices to be inverted are block diagonal.
Therefore, IMEX methods require fewer time steps comparing with fully explicit method, and the calculation for each step is easily parallelizable.  
For a non-relativistic system where the propagating speed of gravity wave and fluid is much slower than the light speed, fully implicit method is a better choice for neutrino transport in CCSN simulation.
But for a relativistic circumstance that we have, where the gravity, fluid, and neutrino have a relativistic propagating speed, IMEX method
balances the efficiency and cost.

@@

>>>>

Core-collapse supernovae (CCSNe) are the explosions of massive stars that end their lives.
CCSNe are a dominant source of heavy elements and play an important role in many astrophysical phenomena, such as neutron star and black hole formation.  
Furthermore, these explosions occur at energies and densities relevant to address fundamental questions in nuclear, particle, and gravitational physics. 
A solid theoretical framework for the CCSN explosion mechanism may help answer important questions in fundamental physics.\cite{janka_etal_2007}

One essential part of the explosion mechanism is neutrino transport.
The neutrino energy deposition drives the CCSN explosion.\cite{mezzacappaMesser_1999}
Ideally neutrino transport would be modeled by the Boltzmann transport equation, which is a integro-partial-differential equation evolving a phase-space distribution function $f$.\cite{Bruenn_1985}
Simulating the neutrino transport implies finding a solution of the Boltzmann equation for a specific domain and period with acceptable accuracy.

However, solving the multi-dimensional Boltzmann transport equation is expensive.
To balance physical fidelity and computational expediency, an approximate method called the two-moment method has been adopted.\cite{mezzacappaMesser_1999}
In the two-moment method, the evolved variables are the zeroth and first angular moments of the distribution function $f$: the spectral particle density $\cJ$ and flux $\bcH$, respectively.
However, the transport term in the Boltzmann transport equation is the dot product of the velocity vector and the gradient of the phase-space distribution function.
Integrating this term with unit direction vector as the weight introduces the second angular moment $\bcK$ to the moment equations.
Knowledge of $\bcK$ is needed to close the equation system.
Therefore, an algebraic closure that gives an approximate value of $\bcK$ given $\cJ$ and $\bcH$ is needed.
The two-moment method has been widely applied with different algebraic closures, such as the Minerbo\cite{minerbo_1978} closure(e.g.~{O'Connor} and {Couch}\cite{oConnorCouch_2018}, Pan and et al\cite{pan_etal_2018}, Glas et al\cite{glas_etal_2018}, and Just et al\cite{just_etal_2018} ) and the Levermore\cite{levermore_1984} closure(e.g.~Vartanyan et al\cite{vartanyan_etal_2018}, Cabezon et al\cite{cabezon_etal_2018}, and Kuroda et al\cite{kuroda_etal_2016}).

Applying the two-moment method alone does not guarantee an affordable solution.
To be precise, neutrinos interact with the background, and the characteristic time scale of these interactions, which can be $\sim\mathcal{O}(10^{-13})$~second, is short compared to the duration of the CCSN explosion, which $\sim\mathcal{O}(1)$~second.  
This makes the time step needed by an explicit time integrator prohibitively small. 
On the other hand, solving the moment equations fully implicitly requires inverting globally many band-structured matrices whose sizes depend on the spatial discretization.
Such a global inversion is both expensive and unfriendly to parallelization.
To circumvent these challenges, implicit-explicit (IMEX) methods are taken into consideration.
By treating the transport terms in the two-moment equations explicitly and the collision terms implicitly, IMEX methods are subject only to a time step governed by the explicit transport terms, and the matrices to be inverted are block diagonal.
Therefore, IMEX methods require fewer time steps comparing with fully explicit method, and the calculation for each step is easily parallelizable.  
For a non-relativistic system where the propagating speed of gravity wave and fluid is much slower than the light speed, fully implicit method is a better choice for neutrino transport in CCSN simulation.\cite{cernohorskyVanWeert_1992}(read ?)
But for a relativistic circumstance that we have, where the gravity, fluid, and neutrino have a relativistic propagating speed, IMEX method
balances the efficiency and cost.

@@

\rc{Reference check}
